---
title: "The Effect of Airbnb's Market Entry on Hotels in Berlin: A Regression Discontinuity Approach"
subtitle: 'Final Paper for Introduction to Collaborative Social Science Data Analysis'
author: "Daniel J Murphy & Paulo H Kalkhake"
date: "December 9, 2016"
output: pdf_document
toc: FALSE
bibliography: Bibliography.bib
linestretch: 1.5
number_sections: TRUE
fig_caption: TRUE
inline_notes: TRUE
csl: international-economic-journal.csl

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = 'pdf')

# Set working directory
try(setwd("/Users/Paulo/GitHub/FinalPaper"),silent=TRUE)
try(setwd("/Users/djm113/Documents/GitHub/FinalPaper"),silent=TRUE)

#loading graphs packages
library(ggplot2)
library(grid)
library(scales)
library(RColorBrewer)
library(stargazer)

# Dynamical Link to Data/Packages R script file (takes some time!)
# source("assignment03_analysis.R")

##create a theme for all graphs (courtesy of Max Wolf)
fte_theme <- function() {
      
# Generate the colors for the chart procedurally with RColorBrewer
palette <- brewer.pal("Greys", n=9)
color.background = palette[2]
color.grid.major = palette[3]
color.axis.text = palette[6]
color.axis.title = palette[7]
color.title = palette[9]
      
# Begin construction of chart
theme_bw(base_size=9) +
        
# Set the entire chart region to a light gray color
theme(panel.background=element_rect(fill=color.background, color=color.background)) +
theme(plot.background=element_rect(fill=color.background, color=color.background)) +
theme(panel.border=element_rect(color=color.background)) +
      
# Format the grid
theme(panel.grid.major=element_line(color=color.grid.major,size=.25)) +
theme(panel.grid.minor=element_blank()) +
theme(axis.ticks=element_blank()) +
      
# Format the legend, but hide by default
theme(legend.position="none") +
theme(legend.background = element_rect(fill=color.background)) +
theme(legend.text = element_text(size=7,color=color.axis.title)) +
      
# Set title and axis labels, and format these and tick marks
theme(plot.title=element_text(color=color.title, size=10, vjust=1.25)) +
theme(axis.text.x=element_text(size=7,color=color.axis.text)) +
theme(axis.text.y=element_text(size=7,color=color.axis.text)) +
theme(axis.title.x=element_text(size=8,color=color.axis.title, vjust=0)) +
theme(axis.title.y=element_text(size=8,color=color.axis.title, vjust=1.25)) +
      
# Plot margins
theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}

```
#Abstract

#Introduction
The effects of Airbnb on the Berlin housing market have recently been the subject of a great deal of scrutiny. In 2014, the city passed a law that forbade renting out entire apartments to tourists on Airbnb. This came in response to an unprecedented 56 percent rise in rents in Berlin between 2009 and 2014, for which Airbnb and other online hosting platforms received some of the blame [@France-Presse:2016aa]. The city's law was widely welcomed by Berliners hoping to slow the forces of gentrification, but it was also welcomed by another Airbnb victim in Berlin; the hotel industry. 

While Berliners fought tourists for apartments, hotels were quietly fighting Airbnb for tourists. Airbnb represents an unprecedented challenge for the long-time incumbents, whose average price of 80 euros per night cannot compete with the average Airbnb price of 55 euros per night [@Alsino-Skowronnek:2015aa]. To make matters worse, the rise of Berlin as the place to be happened to coincide with the rise of Airbnb as the place to stay. This prevented the hotels of Berlin from fully realizing profits that were, in their minds, rightfully theirs. Airbnb, however, claims that their service is complementary, since 77 percent of Berlin's Airbnb listings are outside of the main hotel districts [@Airbnb:2015aa]. 

However, a paper by @zervas2016rise found that the rise of Airbnb had a negative effect on hotel revenue in the state of Texas. This shows that, while Airbnb may be complementary, they also compete with the incumbent hotel industry. In this paper we will seek to illustrate the magnitude of the "Airbnb effect" on hotels in Berlin. 

# Literature Review

## Airbnb Literature
While the literature on the effect of Airbnb on hotels is quite limited, there is one prominent paper that helped inspire this investigation. In their paper, @zervas2016rise used a difference in differences method (DID) to measure the impact that Airbnb's presence had on Texas hotel revenue, and how this impact differed by region and over time. They found that a 10 percent size increase of the Airbnb market in Texas resulted in a .39 percent decrease in hotel revenue. This effect varied widely by region, however. Austin, for example showed a 8 to 10 percent impact on hotel revenue. Zervas et al. also went on to investigate the impact of Airbnb's presence on different types of hotels and how that presence may effect hotel pricing models. They found that hotels with lower prices are more affected than higher priced hotels, and that Airbnb has significantly hindered hotels' ability to raise prices during high demand periods. 

Some literature does exist on Airbnb's entry into Berlin, however much of this is focused on the impact on rents in the city. @schafer2016misuse, for example, found that 5,555 residential apartments are currently being misused for Airbnb in Berlin. However, as the *Zweckentfremdungsverbot* (Airbnb misuse law) came into effect this year, the main focus of this issue is likely to be enforcement going forward. 

More broadly, there is already diverse body of literature available on different aspects of Airbnb. In @zervas2015first, the authors explore user ratings on Airbnb and TripAdvisor. They find that the average user rating of Airbnb is significntly higher than that of the average hotel on TripAdvisor, 4.5 out of 5 stars compared to 3.8 out of 5 stars, respectively. Their paper does not purport to be conclusive in determining the reason for this, but rather takes a first stab at quantifying the difference. They theorize that Airbnb hosts, being more entrepreneurial than their hotelier competitors, go above and beyond to ensure that they recieve good reviews that bolster their individual "brand".

Along similar lines, @gutt2015sharing investigated how Airbnb hosts change their prices in response to their first review. They found that increased "rating visibility" causes hosts to raise their prices by an average of 2.69 euros. 

## Tourism Literature
In terms of our focus, a working paper by @storsmotives indicates that the motives for a tourist to use Airbnb rather than a hotel are both monetary and otherwise. In fact, they found that guests' expectations of having a more authentic, personal experience during their stay were just as important as the monetary aspect. This suggests that the hotel industry in Berlin is not only being undercut in pricing, but also outclassed in the experience they provide.




Based on the extended literature review, two hypotheses will guide our thinking. 

> *H1: The higher the Airbnb supply in a given district in Berlin, the lower the hotel occupancy rate will be in that same district.*

> *H2: Since Airbnb's listings are concentrated in districts with low hotel density, the effect of substitution will be more pronounced in those districts.*

#Data & Variables

Our data comes from three different sources, the Statistical Information System Berlin/Brandenburg (SBB) [@Brandenburg:2016aa], the Federal Statistical Office and the statistical offices of the Länder (FSO) ^[Both databases use JAVA-based website, which did not allow direct web scraping. The data was manually downloaded.] [@Germany:2015aa], and *InsideAirbnb.com* [@Airbnb:2016aa]. In order to conduct our analysis we neede to clean, merge and manipulate these data sets. 
From the Statistical Information System Berlin/Brandenburg, we collected monthly data on the number of overnight stays and the number of guests to arrive at their accommodations in the reporting period in Berlin [@Brandenburg:2016aa]. The surveys are carried out at the beginning of each month and refer to the reporting period of the previous month. The results are organized regionally according to districts and municipalities, allowing us to have specific data for each of the twelve districts in Berlin. We also gathered data for yearly household income groups and the number of employed and unemployed people per district. Based on the data, we calculated a yearly average household income and unemployment rate per district.

From the Regional Database of Germany, we collected data on the supply of hotel beds in each district in a given year. Unfortunately this data is only recorded annually. Hence, our analysis must assume that the number of beds in each hotel stay constant throughout the year. For our main dependent variable, we used data on the number of beds generally available in each district per day in a given year and the monthly data on the number of overnight stays from guests to compute a variable as a proxy for hotel occupancy (as a percentage).

From *InsideAirbnb.com*, we scraped data on 15,368 listings, i.e. apartments or rooms, for Berlin from August, 2008 until October, 2015. This data was extracted from the Airbnb site between 18 July 2015 and 6 January 2016. Amongst 92 variables for each listing covering topics ranging from room price to information on the host, the data includes (1) the neighbourhood of each listing, (2) the date that an Airbnb host signed up, and (3) the date of the last review of each listing. We further include a proxy for Airbnb's market entry. It is based on date of of acquisition of the German peer-to-peer market place Accoleo in June 2011, which is considered Airbnb's first move to expand their business model to the European market (http://thenextweb.com/insider/2011/07/06/airbnb-takes-on-europe-will-it-revolutionize-the-industry-again/). 

Our most significant methodological challenge is the absence of precise listing availability during our period of interest, as it is not directily available in the data. However, in keeping with the @zervas2016rise methodology, we used data on when the host became Airbnb member as a proxy for market entry. We then construct a variable for cumulative supply based on this information, i.e. the total number of listings in a district belonging to users that have signed up on Airbnb in prior to that month. This is not ideal, as this proxy does not take into account whether or not a listing was available in a given month. However, Airbnb itself is also unable to produce exact supply data. This is because owners do not accurately update their listings' availability. 

The final data set covers 720 monthly obversations across tewlve districts in Berlin and covers the time period between 2010 and 2014.

#Methodology & Empirical Anlysis
##Descriptive Analysis
Airbnb`s popularity amongst users and hosts has increased substantially since it was founded in August 2008. As Berlin's popularity has increased over that same period, many Berliners began listing apartments on Airbnb. The growth in Airbnb listings has not been equal across all neighbourhoods in Berlin, but it has been positive in every year from 2010 to 2014. That trend has accelerated each year, with more and more Berliners listing apartments on the site. 

```{r, echo=FALSE}
# Creating end of year subset
NID_Subset <- subset(analysis_simple, analysis_simple$month==12)

ggplot(NID_Subset, aes(x=year, y=AB_supply, group=NID)) +
    geom_line(alpha=0.5, color="#c0392b") +
    fte_theme () +
    labs(x="Year", y="Airbnb Listings", title="Growth of Airbnb Listings in Berlin (2010 - 2014)")


```


Upon plotting Airbnb supply against hotel room occupancy, we were surprised to see a clear positive correlation. However, we realized the importance of accounting for general demand in an area as it becomes more attractive to tourists. @zervas2016rise accomplished this by controlling for passengers listing the local airport as their final destination. We will incorporate a similar statistic in our final analysis, as tourism in Berlin has increased dramatically in recent years. Given confirmation from Zervas et al. that Airbnb does indeed compete with hotels, we believe that this increase in Berlin's popularity is largely responsible for the positive correlation we observe here.

```{r, echo=FALSE}

ggplot(analysis_simple, aes(x=AB_supply, y=occup_rate)) +
    geom_point(alpha=0.5, color="#c0392b") +
    geom_smooth(method = "lm", se = FALSE, color="black", fill="black") +
    fte_theme() +
    labs(x="Airbnb Listings", y="Hotel Occupance Rate", title="Effect of Airbnb Listings on Hotel Occupancy Rates (2010 - 2014")

```


##Inferential Analysis
Thus for further analysis of the data, we included the number of guests per month for each district. 

```{r, echo=FALSE}

simple_model <- lm(occup_rate ~ AB_supply + guests, data=analysis_simple)

```

``` {r, results='asis'}

stargazer(simple_model, header=FALSE, type='latex')

```

Even when including guests in our model we find that Airbnb supply has a highly significant positive effect on hotel occupancy, with a p-value on the .01% level. This surprised us as well, and warrants further exploration of potential explanations in our final analysis. 

Unlike @zervas2016rise, we cannot use hotel revenue as our main unit of analysis due to data limitations. Instead, our analysis will focus on hotel occupancy rates. Also unlike the model used by the authors, our analysis cannot take advantage of a treatment and control region accounting for the different Airbnb market entry patterns. Instead, we propose the use of a Regression Discontinuity Model (RDM) [@jacob2012practical]. 

By introducing a binary variable which is equal to one for all observations after Airbnb's market entry in Berlin and interacting that variable with all regressors in our model, we account for all the changes in hotel occupancy rates before and after Airbnb's market entry in Berlin, i. e. the dicontinuity at market entry. 

Taking the arguments above into account, our main specification for our proposal would look like this:

> (1) ${\log OccupRate}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

Here, log *OccupRate* is the occupancy rate for all hotels in district *i* at time *t*. Log *Abb* is the log of the approximated total number of Airbnb listings available in district *i* at time *t*. Further, we control for economic conditions (unemployment rate and GDP per capita at federal level) and a district-specific linear time trend to account for unobserved heterogeneous variation across districts by using a month-year time dummy *$\tau$*.

#Results

#Discussion & Conclusions

\newpage
##Bibliography

