---
title: "The Effect of Airbnb on Hotels"
subtitle: "Evidence from Berlin"
author: 
  - Paulo H. Kalkhake ^[Matriculation N° 147 153, p.kalkhake@mpp.hertie-school.org]
  - Daniel J. Murphy
date: "15 December 2016"
output: 
  pdf_document: 
linestretch: 1.5
number_sections: TRUE
keep_tex: true
latex_engine: pdflatex
fig_caption: TRUE
inline_notes: TRUE
geometry: top=2cm, right=3cm, left=3cm, bottom=3cm
bibliography: Bibliography.bib
abstract: "This paper analyses the impact of Airbnb’s market entry in June 2011 on the hotel industry in Berlin. By using fixed effects strategies we explore this relationship using district-level data for hotel occupancy and Airbnb listing data from 2010 until 2014. When controlling for economic conditions and increasing air travel to Berlin, we estimate that a one percent increase in Airbnb supply relates to a .01 percentage points decrease in hotel occupancy rates. Moreover, the impact is non-uniformly distributed, with lower-priced hotels and those hotels not catering to business travelers being the most affected. Our work provides empirical evidence that Airbnb may be successfully competing with incumbent firms. Future research should further explore the causal effect of Airbnb market entry in cities. Policy makers should focus on creating a level-playing field for both hotels and Airbnb hosts to create a fair legal framework for all market participants."
documentclass: article
csl: "finalpaper/apa.csl"
---

```{r setup, include=FALSE, fig.height=3.5, fig.width=8}
knitr::opts_chunk$set(echo = FALSE, dev = 'pdf')

# Set working directory
try(setwd("/Users/Paulo/GitHub/FinalPaper"),silent=TRUE)
try(setwd("/Users/djm113/Documents/GitHub/FinalPaper"),silent=TRUE)

#loading graphs packages
library(ggplot2)
library(grid)
library(scales)
library(RColorBrewer)
library(stargazer)
library(captioner)
library(ggplot2)
library(car)
library(plm)
library(lubridate)
library(knitr)
library(pastecs)
library(Hmisc)
library(curl)
library(rio)
library(stringr)
library(plyr)
library(xml2)
library(DataCombine)
library(tidyr)
library(data.table)
library(dplyr)
library(zoo)
library(xtable)
library(texreg)
library(stargazer)
library(foreign)
library(gplots)
library(lubridate)

```

```{r include=FALSE, cache=TRUE}

# Dynamical Link to Data/Packages R script file (takes some time!)
source("code/finalpaper_analysis.R")

```

\begin{centering}
\vspace*{\fill}
  \pagenumbering{gobble}
\vspace*{\fill}
\end{centering}


\newpage
  \pagenumbering{arabic}
  \setcounter{page}{2}
  \setlength{\parindent}{0.25in}
\bibliographystyle{apacite}

##Introduction

Airbnb is a young company that has revolutionized the accomodation industry in much the same way that Uber has revolutionized transportation. It is an online platform that allows "hosts" to share their a room in their home or a rental home with travelers. For hosts, using Airbnb can be a valuable source of extra income. For guests, Airbnb provides accomodations that allows travelers to both save money and have a more authentic experience during their stay. Airbnb guests often remark that they interacted with their hosts and felt more like a local than a tourist. Indeed, Airbnb's recent "Don't Go There, Live There" advertising campaign is a reflection of their philosophy in reshaping the accomodation industry. 

In Berlin more than 20,000 Airbnb hosts have been active over the past year, hosting more than half a million guests. According to Airbnb, their presence in Berlin brings more economic activity to the city, particularly to neighborhoods outside of the main hotel districts. Yet, Airbnb's entry in Berlin has been highly controversial. In 2014, the city passed a law that forbade renting out entire apartments to tourists on Airbnb. This came in response to an unprecedented 56 percent rise in rents in Berlin between 2009 and 2014, for which Airbnb and other online hosting platforms received some of the blame [@France-Presse:2016aa]. The city's law was widely welcomed by Berliners hoping to slow the forces of gentrification, but it was also welcomed by another Airbnb victim in Berlin; the hotel industry.

While Berliners compete with tourists for apartments, the hotel industry has been quietly fighting a battle of their own against Airbnb. Airbnb represents an unprecedented challenge for the long-time incumbents, whose average price of 80 euros per night in Berlin cannot compete with the average Airbnb price of 55 euros per night [@Alsino-Skowronnek:2015aa]. Moreover, Airbnb enjoys some real advantages over the hotel industry due to its new, still largely unregulated business model. For example, its informal business model has thus far allowed it to avoid the taxes and regulations that hotels and other traditional accomodation providers have to deal with [@guttentag2015airbnb]. For their part, Airbnb claims that their service is complementary to hotels and that most of their listings are outside of cities' main hotel districts [@Airbnb:2015aa]. Airbnb guests are also more likely to stay at the host's property long term than a hotel guest is, and it is not uncommon to find Airbnb listings that are rented out to the same guest for a month or longer. Yet despite these differences, many believe there must be some "Airbnb effect" on the hotel industry. A paper by @zervas2016rise found that the rise of Airbnb had a negative effect on hotel revenue in the state of Texas. This shows that, while Airbnb may be complementary, they also compete with the incumbent hotel industry. 

In this paper we will seek to illustrate the magnitude of the "Airbnb effect" on hotels in Berlin. To that end, this paper will proceed as follows. First, we will review the relevant literature in the field and outline our hypothesis going forward. Second, we will discuss the data sources and variables that we will use to explore this relationship. Third, we will outline the methology we use in this study and describe our model in detail. Fourth, we will analyze the data both descriptively and inferentially to evaluate our hypotheses. Finally, we will discuss our results, the policy implications of our findings, and opportunities for future research. 

##Literature Review
While the literature on the effect of Airbnb on hotels is quite limited, there is a diverse body of literature available on different aspects of Airbnb's business. In Berlin, much of the discussion has focused on the impact of Airbnb on rents in the city. @schafer2016misuse, for example, found that 5,555 residential apartments are currently being misused for Airbnb in Berlin. However, as the *Zweckentfremdungsverbot* (Airbnb misuse law) came into effect this year, the main focus of this issue is likely to be enforcement going forward. 

Like Uber, Airbnb has also been widely praised for providing a more personalized experience for consumers in an industry that may have lacked it before. In both @yannopoulou2013user and @guttentag2015airbnb, the authors find that the desire for a personalized user experience played a role in the rise of Airbnb. This new expectation has changed the very nature of the accomodation and tourism industry. A working paper by @storsmotives shows that this trend extends to Berlin, and indicates that the motives for a tourist to use Airbnb rather than a hotel are both monetary and otherwise. In fact, Stors & Kagermeier found that guests' expectations of having a more authentic, personal experience during their stay were just as important as their monetary concerns. This suggests that the hotel industry in Berlin is not only being undercut in pricing, but also outclassed in the experience they provide.

In @zervas2015first, the authors explore user ratings on Airbnb and TripAdvisor. They find that the average user rating of Airbnb is significantly higher than that of the average hotel on TripAdvisor, 4.5 out of 5 stars compared to 3.8 out of 5 stars, respectively. They theorize that Airbnb hosts, being more entrepreneurial than their hotelier competitors, go above and beyond to ensure that they receive good reviews that bolster their individual "brand". Along similar lines, @gutt2015sharing investigated how Airbnb hosts change their prices in response to their first review. They found that increased "rating visibility" causes hosts to raise their prices by an average of 2.69 euros. 

In @guttentag2015airbnb, the author argues that while many points discounting Airbnb's ability to compete with the traditional accomodation industry are valid, there is still no reason to completely discount the young disruptor's impact. Airbnb will probably never completely replace the hotel industry, but it is already large enough to have an impact nevertheless. Indeed, Guttentag estimates that Airbnb sold about 15 million room nights in 2012, which would have made its footprint on the industry that year similar to that of Fairfield Inn & Suites or InterContinental, two major hotel brands. Guttentag concludes his paper with some recommendations for future research on Airbnb, and highlights the need to investigate the impact of Airbnb on hotel occupancy rates and room prices. 

Along those lines, there is one prominent paper that helped inspire our investigation. In their paper, @zervas2016rise used a difference in differences method (DID) to measure the impact that Airbnb's presence had on Texas hotel revenue, and how this impact differed by region and over time. They found that a 10 percent size increase of the Airbnb market in Texas resulted in a .39 percent decrease in hotel revenue. This effect varied widely by region, however. Austin, for example showed a 8 to 10 percent impact on hotel revenue. Zervas et al. also went on to investigate the impact of Airbnb's presence on different types of hotels and how that presence may effect hotel pricing models. They found that hotels with lower prices are more affected than higher priced hotels, and that Airbnb has significantly hindered hotels' ability to raise prices during high demand periods.

In order to add to this body of research, and at the behest of @guttentag2015airbnb, our investigation will focus on the effect of Airbnb listing on hotels in Berlin. Specifically, the following hypothesis will guide our thinking. Controlling for district and economic conditions, we assume that *the higher the Airbnb supply in Berlin, the lower the hotel occupancy rate will be.*

# Data & Variables
To approach our research question we use data  from four different sources, the Statistical Information System Berlin/Brandenburg (SBB) [@Brandenburg:2016aa], the Federal Statistical Office and the statistical offices of the Länder (FSO) ^[Both databases use JAVA-based website, which did not allow direct web scraping. The data was manually downloaded.] [@Germany:2015aa], *InsideAirbnb.com* [@Airbnb:2016aa], and Eurostat. In order to conduct our analysis we needed to clean, merge and manipulate these data sets. 

The surveys from the Statistical Information System Berlin/Brandenburg are carried out at the beginning of each month and refer to the reporting period of the previous month. The results are organized regionally according to districts and municipalities, allowing us to use specific data for hotels in each of the twelve districts in Berlin. From this survey data, we collected monthly data on the number of overnight hotel stays, the number of hotel beds in each district, and the number of guests to arrive at their accommodations in the reporting period in Berlin [@Brandenburg:2016aa].  

From the Federal Statistical Office and the statistical offices of the Länder (FSO), we gathered data for yearly household income groups and the number of employed and unemployed people per district. Based on the data, we calculated a yearly average household income and unemployment rate per district [@Germany:2015aa].

From *InsideAirbnb.com*, we scraped data on 15,368 listings, i.e. apartments or rooms, for Berlin from August, 2008 until October, 2015. This data was extracted from the Airbnb site between 18 July 2015 and 6 January 2016. This data set contained 92 variables for each listing covering topics ranging from room price to information on the host. For our purposes, the key variables were (1) the neighbourhood of each listing, (2) the date that an Airbnb host signed up, and (3) the date of the first review of each listing [@Airbnb:2016aa].

From Eurostat [-@yearbook2015eurostat], we gathered monthly data on the number of passengers to arrive in Berlin's two major airports, Tegel and Schonefeld. We use this data to construct a variable that combines the arrivals to both airports every month. In keeping with the @zervas2016rise methodology, we use this variable as our control for the dramatic increase in the popularity of Berlin. 

Our analysis also includes a proxy for Airbnb's market entry. It is based on date of of acquisition of the German peer-to-peer market place Accoleo in June 2011, which is considered Airbnb's first move to expand their business model to the European market [@TNW:2011aa]. 

Our most significant methodological challenge was the absence of precise listing availability during our period of interest, as it is not directly available in the data. This would have been the ideal data with which to construct our Airbnb supply variable. However, Airbnb itself is also unable to produce this type of precise data, as hosts do not accurately update their listings' availability. In keeping with the @zervas2016rise methodology, we countered this problem by using review data as a proxy for availability. For both our cumulative and dynamic Airbnb supply variables, we calculated the date 6 months prior to a listing's first review as its listing date. If there was no first review, we used the date that a host signed up on the site as the listing date. If a listing had neither, it was excluded from the analysis. Only five listings were excluded on this basis. 

Our analysis includes two measures of district-specific Airbnb supply, a cumulative measure and a dynamic measure. The cumulative measure simply adds new listings to the supply either as they either recieve first reviews or as their owner registers on the site. The weakness of this measure is that it does not eliminate listings that may have ceased to be available. To account for this, we constructed a dynamic measure of Airbnb supply. Like the cumulative measure, the dynamic variable uses either the first review (minus 6 months) or the hosts sign-up date as the listing date. From there, the dynamic measure searches the six months prior to each month in our analysis to find a review. If it does not find one, the listing is assumed to have become unavailable, and drops out of the supply. However, when a listing receives a new review it reenters the dynamic supply. 

Admittedly, our approach of using Airbnb reviews as a proxy for Airbnb supply produces something of a conservative Airbnb supply estimate. @fradkin2015bias found that only 67% of Airbnb guests leave a review following their stay. However, in the absence of exact listing availability data the review data must suffice. 

Our dependent variable throughout our analysis will be the occupancy rate of hotels in a given district. We could also have used the number of overnight stays as our dependent variable, as it was directly available in the data. However we did not believe this would sufficiently account for the growing supply of hotel beds in Berlin each year. Therefore, we calculated district specific hotel occupancy rates for each month of our analysis by dividing the number of overnight stays in that month by the product of the supply of hotel beds in each district multiplied by the number of days in each month. (growth rate: ?)

$${Occupancy Rate}_{it} =  \frac {{Overnight Stays}_{it}}{{Hotel Beds}_{it}*{days}_{t}}$$

The final data set covers 720 monthly obversations across twelve districts in Berlin and covers the time period between 2010 and 2014.

## Analysis
### Descriptive Analysis
Airbnb`s popularity in Berlin has significantly increased since it's official debut in the city in June 2011. Flatsharing was already on the rise in the ciy, but Airbnb's acquisition of Accoleo consolidated the Berlin flatsharing industry into one well-known name. As Berlin's increase in popularity coincided with the rise of Airbnb, many Berliners used the site to list apartments and rooms. The growth in Airbnb listings has not been equal across all neighbourhoods in Berlin, but it has dramatically increased since 2010. 

In Figure 1, we can see the growth in Cumulative Airbnb Supply in each neighbourhood from 2010 to 2014. Each of the most popular neighbourhoods (Friedrichshain-Kreuzberg, Mitte, Pankow, & Neukoelln) is shown to have more than 1,500 listings by 2014. Growth is shown to accelerate in the city's most popular disticts each year, and Friedichshain-Kreuzberg's total number of listings is well above 2,500 by the end of 2014. 

```{r, echo=FALSE, fig.cap='Cumulative Airbnb Supply per Neighbourhood (2010 - 2014)'}

analysis_simple$ymym <- as.yearmon(analysis_simple$year_month)

ggplot(analysis_simple, aes(x=ymym, y=AB_supply, group=neighbourhood, colour=neighbourhood)) +
    geom_line(alpha=0.7) + 
    labs(x="Months", y="Number of Airbnb Listings") + 
    scale_colour_discrete(name ="Neighbourhood") +
    scale_x_yearmon(as.yearmon(analysis_simple$ymym), format = "%Y-%m", n = 6)

```

\newpage

However in Figure 2, we can see that our Dynamic Airbnb Supply tells a slightly different story. In mid-2014 we see Airbnb supply peak in each district before falling dramatically in the second half of the year. Indeed, Friedrichshain-Kreuzberg's supply even appears to drop below it's supply level at the start of the year. Upon closer examination, the supply peak appears to coincide with the passage of the *Zweckentfremdungsverbot* (Airbnb misuse law) in the summer of 2014. This highly publicized law prohibited listing full apartments on Airbnb in Berlin, but allowed the continued listing of individual rooms. However, the law did not come into effect until 2016, and *InsideAirbnb.com* data from 2015 shows that Airbnb supply came back stonger than ever immediately following our obervation period. These considerations led us to conclude that it was better to use our Cumulative Airbnb Supply variable for our inferential analysis. Despite the shortcomings of the cumulative variable, it more accurately represents the overall supply growth of Airbnb in Berlin, which both variables likely understate anyway due to the 33 percent of Airbnb guests who do not leave reviews [@fradkin2015bias].

```{r, echo=FALSE, fig.cap='Dynamic Airbnb Supply per Neighbourhood (2010 - 2014)'}

danalysis_simple$ymym <- as.yearmon(danalysis_simple$year_month)

ggplot(danalysis_simple, aes(x=ymym, y=AB_supply, group=neighbourhood, colour=neighbourhood)) +
    geom_line(alpha=0.7) + 
    scale_x_yearmon(as.yearmon(analysis_simple$ymym), format = "%Y-%m", n = 6) +
    labs(x="Months", y="Number of Airbnb Listings") + 
    scale_colour_discrete(name ="Neighbourhood") 

```

In order to take a first look at the relationship between Airbnb supply and hotel occupancy rates, we plotted the log of Airbnb supply against hotel occupancy rates. We chose to work with the log of Airbnb supply here in order to normalize the distribution, which was otherwise quite skewed. We refrained from taking the log of hotel occupancy rates since they are smaller than one, and appear to have a fairly normal distribution. 
\newpage

Upon plotting this relationship, we were surprised to see a clear positive correlation and a correlation coeffecient of 0.343 *(cf. Figure 3)*. We believe this counterintuitive result is likely due to an omitted variable that is driving hotel occupancy rates. Given the increase in Berlin's popularity, we identified increased accomodation demand in general as the most likely suspect. @zervas2016rise controlled for this effect by using the number of passengers listing the local airport as their final destination. We replicated this approach by using Eurostat data on the number of passengers arriving at Berlin's two major airports. Upon cleaning this data, we see an increase of _____________ in the number of passengers arriving in Berlin. Additionally, our inferential analysis will incorporate a time specific trend to account for unobserved heterogeneity. Given confirmation from Zervas et al. that Airbnb does indeed compete with hotels, we believe that incorporating the proper controls into our model will eliminate the positive correlation we observe here.

```{r, echo=FALSE, fig.cap='Effect of Increase in Airbnb Listings on Berlin Hotel Occupancy Rates (2010 - 2014)'}

marketentry_dynamic <- subset(danalysis_simple, danalysis_simple$dmarketentry==1)
marketentry_static <- subset(analysis_simple, analysis_simple$dmarketentry==1)

ggplot(marketentry_dynamic, aes(x=log_ABsupply, y=occup_rate)) +
    geom_point(alpha=0.5, color="#c0392b") +
    geom_smooth(method = "lm", se = FALSE, color="black", fill="black") +
    labs(x="Log Number of Airbnb Listings", y="Hotel Occupance Rate")

```

###Methodology
In order to explore the relationship between Airbnb supply and hotel occupancy rates further, we apply several different approaches of multivariate analysis. After beginning with an OLS approach, we use fixed effects models to account for unobserved time-varying and neighbourhood-specific factors. These are constant over time and correlated with the control variables, independent variables, and the hotel occupancy rate. In doing so, we do not systematically vary both between different cities and over time, because the models look at neighbourhood-specific within-variation [@wooldridge2015introductory]. 

As discussed above, our analysis differs from that of @zervas2016rise in that our unit of analysis is hotel occupancy rates rather than hotel revenue. However our methodology differs from Zervas as well, as our analysis cannot take advantage of a treatment and control regions accounting for different Airbnb market entry patterns. Instead, we use regional fixed effects to account for time-invariant differences in hotel occupancies in each region. 

Considering these arguments, the main specification for our analysis looks like this:

$${OccupRate}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$$

Here, *OccupRate* is the occupancy rate for all hotels in district *i* at time *t*. Log *Abb* is the log of the approximated total number of Airbnb listings available in district *i* at time *t*. Further, we control for economic conditions (unemployment rate and GDP per capita at federal level) and a district-specific linear time trend to account for unobserved heterogeneous variation across districts by using a month-year time dummy *$\tau$*.

We ran a Hausman test in order to decide whether to use a fixed or random effects model, since a regular OLS regression does not consider heterogeneity across groups or time (SOURCE: Green, 2008). This test checks to see if the unique erros are or are not correlated with the regressors. Upon execution, the test showed a p-value of 1 (check whether this is really true). Thus, we reject the null hypothesis that they are not. The Breusch-Pagan Lagrange multiplier test confirms our results, as there is a panel effect, i.e. a significant difference across neighbourhoods (p-value = 0.01).

#Results & Discussion
Table 1 compares the results for the different models. 

``` {r, echo = FALSE, include=TRUE, message = FALSE, error = FALSE, results = "asis"}

LMII <- function(x, y, z) { 
  (lm(x ~ 0 + z + log_inc + ue_rate + as.factor(NID) + factor_ym, data=y))
}

FEIa <- function(x, y, z) { 
  plm(x ~ z + log_inc + ue_rate, data=y, index=c("NID", "factor_ym"), model="within")
}

FEI <- function(x, y, z) { 
  plm(x ~ z + log_inc + ue_rate + arrivals, data=y, index=c("NID", "factor_ym"), model="within")
}

FEIIm <- function(x, y, z) {
  plm(x ~ z + log_inc + ue_rate + arrivals + marketentry, data=y, index=c("NID", "factor_ym"), model="within")
}

ModelI <- LMII(analysis_simple$occup_rate, analysis_simple, analysis_simple$log_ABsupply)

ModelIa <- FEIa(analysis_simple$occup_rate, analysis_simple, analysis_simple$log_ABsupply)

ModelII <- FEI(analysis_simple$occup_rate, analysis_simple, analysis_simple$log_ABsupply)

ModelIII <- FEIIm(analysis_simple$occup_rate, analysis_simple, analysis_simple$log_ABsupply)

stargazer(ModelI, ModelIa, ModelII, ModelIII,
          digits = 3, 
          covariate.labels = c("Log Airbnb Listings", "Average HH Income (Log)", "Unemployment Rate", "Incoming Passengers", "Market Entry"),
          title ="TITLE BIATCH"
          omit = c("NID", "factor_ym"),  
          column.labels   = c("LM (1)", "FE (2)", "FE (3)", "FE (4)"),
          model.numbers = FALSE,
          omit.stat = c('f', 'ser'), # to nicely fits on the page
          omit.labels = c("Neighbourhood-specific trend", "Time trend"), 
          dep.var.labels = c("Occupancy Rate"),
          out.header = FALSE, 
          model.names = FALSE,
          header = FALSE,
          no.space = TRUE, 
          font.size = "small"
          add.lines = list(c('District/Time FE?', c('No', "Yes", "Yes", "Yes"))),
          type = "latex")

```

Column 1 is a simple Ordinary Least Squares (OLS) model that only includes our economic control variables (neighbourhood specific unemployment rate and average household income) in addition to neighborhood and time specific trends. As stated earlier, we interpret the coefficient *$\beta$* of *$\log Airbnb supply$* as indicative of the extent to which Airbnb listings act as a substitute for hotel rooms. In Column 1 we estimate the coefficient *$\beta$* = 0.009 as being highly significant (p < 0.01). This means that a 10 percent increase in Airbnb listings is associated with a statistically significant 0.09 percentage point increase in the monthly hotel occupancy rate. This positive effect seems counter-intuitive, leading us to beleive that the simple linear model does not adequately analyze the relationship. 

In Column 2 we employ a fixed effects model to account for changes in time and variation across neighbourhoods. Here, we once again find a highly significant (p < 0.01) positive coefficient of 0.02, indicating that a 10 percent increase in Airbnb listings is associated with a statistically significant 0.2 percentage point increase in the monthly hotel occupancy rate. This result shows that we had underestimated the effect of changing from a linear model to fixed effects, and continued to surprise us by revealing a positive relationship between *$\log Airbnb supply$* and monthly hotel occupany rates. However, this model does not control for the dramatic increase in popularity that Berlin has experienced during this time period. 

To control for the effect of Berlins rising popularity, Column 3 follows the methodology of @Zervas2016rise by using data on passengers arriving in Berlin as a proxy for the increasing popularity of the city. Upon adding this control to the fixed effects model, we observe an inverse relationship between *$\log Airbnb supply$* and monthly hotel occupany rates. This effect is highly significant (p < 0.01) with a coefficient of -0.01, meaning that a 10 percent increase in Airbnb listings is associated with a statistically significant 0.1 percentage point decrease in the monthly hotel occupancy rate. This was the relationship we had been expecting to find in our hypothesis, and demonstrates that to a very small degree, Airbnb acts as a substitute for hotels in Berlin. The effect of incoming passengers is weaker still, but positive and highly significant at the p < 0.01 level. This supports our thinking that the increasing popularity of the city has effected hotel occupancy rates. Column 3 also shows an increase in the significance of average household income, which is now significant at the p < 0.01 level. 

Column 4 adds a final control to our analysis, a dummy variable for Airbnb's Berlin market entry. While our observation period is from 2010 to 2014, Airbnb did not officially enter the Berlin market until June 2011. However it seems that the official market entry of the site did not have any significant effect on hotel occupancy rates. The effect of *$\log Airbnb supply$* remains negative and significant at the p < 0.01 level here, albeit with a slightly weakened coefficient (-0.008). 

While our strongest finding, a coefficient of -0.01, is highly significant at the p < 0.01 level, the magnitude of the effect on hotel occupancy rates is very weak. Column 3 indicates that a 10% increase in Airbnb listings is only associated with a 0.1 percentage point decrease in hotel occupancy rates. To put that in perspective, it means that adding about 1,530 Airbnb listings to the Airbnb supply merely causes a 0.1 percentage point decrease in hotel occupancy rates. This suggests what while Airbnb may be affecting the hotel industry in Berlin, it's effect is so small that the hotel industry is unlikely to notice. That is especially truen in light of Berlin's increasing popularity, which seems to be more than making up for any "Airbnb effect" on hotels in Berlin.

``` {r, echo = FALSE, include=TRUE, message = FALSE, error = FALSE, results = "asis"}
#Introduces a binary variable for Airbnb's official market entry + nonlinear
FEV <- function(x, y) {
  plm(x ~ AB_supply + AB_supply_2 + log_inc + ue_rate + arrivals + marketentry, data=y, index=c("NID", "factor_ym"), model="within")
}

ModelIV <- FEV(analysis_simple$occup_rate, analysis_simple, analysis_simple$log_ABsupply)

stargazer(ModelIV, 
          digits = 3, 
          covariate.labels = c("Airbnb Listings", "Sqrd. Airbnb Listings", "Average HH Income (Log)", "Unemployment Rate", "Incoming Passengers", "Market Entry"), 
          omit.stat = c('f', 'ser'), # to nicely fits on the page
          dep.var.labels = c("Occupancy Rate"),
          out.header = FALSE, 
          model.names = FALSE,
          header = FALSE,
          add.lines = list(c('District/Time FE?', rep('Yes', 4))),
          align = FALSE,
          type = "html")

```

Our new main specification takes further into account that the effect of Airbnb is likely best observed, once it has reached a critical threshold or market penetration. However, even when introducing an interaction between log Airbnb supply and the absolute level of Airbnb listings we find that Airbnb supply has a highly significant positive effect on hotel occupancy, with a p-value on the .01 % level. 

Here, we would try to account for the fact that Airbnb supply is more complicated than we have so far defined it. In light of Zervas, we introduce review data, and redefine airbnb supply to include our best inference as to whether Airbnb was available at a given time based off of those reviews. To do that, we would define supply as the number of airbnb's available in a given apartment in a given month. This would be calculated by looking at the date of the first review of the apartment on Airbnb and the date of the individual reviews. Upon the sign up date, the listing would be assigned a "life period" of 6 months, which would then be renewed every month upon each subsequent review. If a listing exceeds its life period without a review, then we assume that it is no longer available. However, once it recieves another review it would then be considered available again for another 6 months. 

Further methods of data exploration

> We then want to control for unobservables that evolve following some trend that can differ from a region to another. This is often done in DD papers by including a linear or a quadratic region-specific time trend. In Zervas et al. (2014), they only report estimates with the quadratic city-specific trend since their results were apparently not really affected by this choice. 

how listing’s characteristics matter?

Airbnb lists all sort of accommodations, from a house or an apartment to shared rooms. This has an influence on how closely they can replace a hotel room, and thus have a different impact on hotel occupancy. We thus excluded all listings categorized as ”shared room”.

In contrast to Zervas et al. (2014) we did not find that Airbnb had a significant negative impact on hotel revenue per available room when looking at Berlin. Several reasons migh

First of all, they had used individual data for all hotels while we only had a computed value on hotel occupancy at the district level. Thus, we measure the impact of Airbnb supply on a much smaller scale and might have some sort of omitted variable bias. 

#Policy Implications & Conclusion
The purpose of this paper was to measure the impact of Airbnb on the hotel industry in Berlin. Whilst a paper by @zervas2016rise found that Airbnb has a negative impact on the hotel industry in Texas, we could not find a similar effect in Berlin. Much to our surprise, we found a positive highly significant effect of Airbnb supply on both, overnight stays and hotel occupancy rates accross all districts between 2010 and 2014. 

From our point of view, this might have different reasons. First, our observation period is not ideal. Aside from its obvious time limitation of only four years, our analysis does not include the year 2015. This is a significant loss for the Airbnb supply variable, as it appears that 2015 was the year that Airbnb supply in Berlin truly skyrocketed. We can see in the raw data from *InsideAirbnb.com*, for example, that the number of listings in Friedricshain-Kreuzberg increased by 342 in March of 2015, compared to 97 in March of 2014. Unfortunately, data limitations in the rest of our variables for the year 2015 prevented us from fully utilizing this Airbnb supply data. 

Second, Airbnb accomodations and hotel rooms may be better considered as complements and not substitutes. 

Perhaps the most likely explanation, however, is that Zervas et al. was simply blessed with much richer data than we were. Specifically with respect to Airbnb supply, we believe that both our investigation and Zervas's would benefit from detailed listing availability data. 

richness using data on hotel level (less aggregated) 

\newpage
#Bibliography
