---
title: "Airbnb vs. Hotels in Berlin: A Regression Discontinuity Approach"
author: "Daniel J Murphy & Paulo H Kalkhake"
date: "2 December 2016"
output: 
  ioslides_presentation:
    #css: https://maxcdn.bootstrapcdn.com/font-awesome/4.4.0/css/font-awesome.min.css
    logo: 
  beamer_presentation: default
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Set working directory
try(setwd("/Users/Paulo/GitHub/FinalPaper/presentation"),silent=TRUE)
try(setwd("/Users/djm113/Documents/GitHub/FinalPaper/presentation"),silent=TRUE)

```

## Airbnb in Berlin

- According to Airbnb, more than 20,000 Berliners have hosted travelers via Airbnb in the past year. 

- In that same period, more than half a million guests have used Airbnb in Berlin.

- Airbnb also reports that the average Berlin host's rating is 4.6 out of 5 stars, and that staying with Airbnb makes guests more likely to return to Berlin. 

Graphic?

## Recent Literaure

- *Rents and Zweckentfremdungsverbot:* Schafer et al found that 5,555 residential apartments in Berlin are currently being misused for Airbnb. 

- *But What About The Hotel Industry?* Guttentag estimates that Airbnb sold about 15 million room nights in 2012. Similar to that of Fairfield Inn & Suites or InterContinental, two major hotel brands.

<<<<<<< HEAD
## Estimation Results

``` {r, echo = FALSE, include=TRUE, message = FALSE, error = FALSE, results = "asis"}

analysis_simple$logABAB <- analysis_simple$AB_supply*analysis_simple$log_ABsupply

ModelIILM <- lm(log_nights ~ log_ABsupply + AB_supply + log_inc + ue_rate + marketentry + NID + factor_ym, data=analysis_simple)

ModelII <- plm(log_nights ~ log_ABsupply + AB_supply + log_inc + ue_rate + marketentry, data=analysis_simple, index=c("NID", "factor_ym"), model="within")

#Model C: introduces an interaction between log_airbnb supply and the absolute level of airbnb supply
ModelC <- plm(occup_rate ~ log_ABsupply + AB_supply + logABAB + log_inc + ue_rate, data=analysis_simple, index=c("NID", "factor_ym"), model="within")

#Model D: introduces a binary variable for Airbnb's official market entry
ModelD <- plm(occup_rate ~ log_ABsupply + AB_supply + logABAB + log_inc + ue_rate + marketentry, data=analysis_simple, index=c("NID", "factor_ym"), model="within")

stargazer(ModelIILM, ModelII, ModelC, ModelD, 
          title="Effect of Airbnb Supply on Hotels in Berlin (2010 - 2014)", 
          digits = 2, 
          covariate.labels = c("log Airbnb listings", "Airbnb listings", "Log*Level", "Average HH income (log)", "Unemployment rate"), 
          omit = c("NID", "factor_ym"),           
          omit.stat = c('f', 'ser'), # to nicely fits on the page
          omit.labels = c("Neighbourhood-specific trend", "Time trend"), 
          dep.var.labels = c("Overnight", "Occupancy Rate"),
          no.space = TRUE,
          out.header = FALSE, 
          model.names = FALSE,
          header = FALSE,
          font.size = 'small',
          add.lines = list(c('District FE?', rep('No', 4))),
          align = FALSE,
          type = "html")

```
=======
- *Airbnb Messes with Texas:* Zervas et al. found that a 10 percent size increase of the Airbnb market in Texas resulted in a .39 percent decrease in hotel revenue.

## Hypotheses

- *H1:* The higher the Airbnb supply in a given district in Berlin, the greater the negative effect on the hotel industry in that same district.

- *H2:* The Airbnb effect is likely the strongest once it has reached a critical mass.

- *H3:* Since Airbnb's listings are concentrated in districts with low hotel density, the effect of substitution will be more pronounced in those districts.

## Data & Variables

We collected data from three separate sources, the Statistical Information System Berlin/Brandenburg (SBB), the Federal Statistical Office and the statistical offices of the Länder (FSO), and *InsideAirbnb.com*. 
>>>>>>> origin/master

Our analysis will focus on the following variables derived from this data. 

- *Occupancy Rate:* The number supply of beds in a given district divided by the number of overnight stays in that district in a given month. 

- *Airbnb Supply:* The number of listings in each neighborhood in any given month whose host has registered on Airbnb.  

## Exploring Our Data

```{r pressure}
plot(pressure)
```

## Building the Model

In order to test our hypotheses, we will employ several models

> (1) OLS Using Overnights as the Dependent Variable
${\log overnights}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

> (2) Fixed Effects with Income and Unemployment as Economic Control Variables
${\log overnights}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

> (3) Fixed Effects with Income, Unemployment, and a Neighborhood Specific Time Trend
${\log overnights}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

> (4) Uses Hotel Occupancy Rate as the Dependent Variable & Introduces an Airbnb Interaction Term. 
${OccupRate}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

## Results


## Results & Policy Implications

Policy Implications

> (1) Unlike Zervas et al., we did not find a significant negative effect of Airbnb on the hotel industry. 

> (2) 


