---
title: "Airbnb vs. Hotels in Berlin: A Regression Discontinuity Approach"
author: "Daniel J Murphy & Paulo H Kalkhake"
date: "2 December 2016"
output: 
  ioslides_presentation:
    #css: https://maxcdn.bootstrapcdn.com/font-awesome/4.4.0/css/font-awesome.min.css
    logo: 
  beamer_presentation: default
#bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


```

## Airbnb in Berlin

- According to Airbnb, more than 20,000 Berliners have hosted travelers via Airbnb in the past year. 

- In that same period, more than half a million guests have used Airbnb in Berlin.

- Airbnb also reports that the average Berlin host's rating is 4.6 out of 5 stars, and that staying with Airbnb makes guests more likely to return to Berlin. 

Graphic?

## Recent Literaure

- *Rents and Zweckentfremdungsverbot:* Schafer et al found that 5,555 residential apartments in Berlin are currently being misused for Airbnb. 

- *But What About The Hotel Industry?* Guttentag estimates that Airbnb sold about 15 million room nights in 2012. Similar to that of Fairfield Inn & Suites or InterContinental, two major hotel brands.

- *Airbnb Messes with Texas:* Zervas et al. found that a 10 percent size increase of the Airbnb market in Texas resulted in a .39 percent decrease in hotel revenue.

## Hypotheses

- *H1:* The higher the Airbnb supply in a given district in Berlin, the greater the negative effect on the hotel industry in that same district.

- *H2:* The Airbnb effect is likely the strongest once it has reached a critical mass.

- *H3:* Since Airbnb's listings are concentrated in districts with low hotel density, the effect of substitution will be more pronounced in those districts.

## Data & Variables

We collected data from three separate sources, the Statistical Information System Berlin/Brandenburg (SBB), the Federal Statistical Office and the statistical offices of the Länder (FSO), and *InsideAirbnb.com*. 

Our analysis will focus on the following variables derived from this data. 

- *Occupancy Rate:* The number supply of beds in a given district divided by the number of overnight stays in that district in a given month. 

- *Airbnb Supply:* The number of listings in each neighborhood in any given month whose host has registered on Airbnb.  

## Exploring Our Data

```{r pressure}
plot(pressure)
```

## Building the Model

In order to test our hypotheses, we will employ a Regression Discontinuity Model...

> (1) ${\log overnights}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

> (2) ${OccupRate}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

## Results


## Results & Policy Implications


