---
title: "The Airbnb Effect & The Berlin Hotel Industry: A Regression Discontinuity Approach"
author: "Daniel J Murphy & Paulo H Kalkhake"
date: "2 December 2016"
output: 
  ioslides_presentation:
    #css: https://maxcdn.bootstrapcdn.com/font-awesome/4.4.0/css/font-awesome.min.css
    logo: 
  beamer_presentation: default
#bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Airbnb in Berlin

- According to Airbnb, more than 20,000 Berliners have hosted travelers via Airbnb in the past year. 

- In that same period, more than half a million guests have used Airbnb in Berlin.

- Airbnb also reports that the average Berlin host's rating is 4.6 out of 5 stars, and that staying with Airbnb makes guests more likely to return to Berlin. 

Graphic?

## Recent Literaure

- *Rents and Zweckentfremdungsverbot:* Schafer et al found that 5,555 residential apartments in Berlin are currently being misused for Airbnb. 

- *But What About The Hotel Industry?* Guttentag estimates that Airbnb sold about 15 million room nights in 2012. Similar to that of Fairfield Inn & Suites or InterContinental, two major hotel brands.

- *Airbnb Messes with Texas:* Zervas et al. found that a 10 percent size increase of the Airbnb market in Texas resulted in a .39 percent decrease in hotel revenue.

## Hypotheses

- *H1:* The higher the Airbnb supply in a given district in Berlin, the greater the negative effect on the hotel industry in that same district.

- *H2:* The Airbnb effect is likely the strongest once it has reached a critical mass.

- *H3:* Since Airbnb's listings are concentrated in districts with low hotel density, the effect of substitution will be more pronounced in those districts.

## Data & Variables

- We collected data from three separate sources, the Statistical Information System Berlin/Brandenburg (SBB), the Federal Statistical Office and the statistical offices of the Länder (FSO), and *InsideAirbnb.com*. 

- *Occupancy Rate:*

- *Airbnb Supply:*

## Slide with Plot

```{r pressure}
plot(pressure)
```

