---
title: "The Effect of Airbnb's Market Entry on Hotels in Berlin: A Regression Discontinuity Approach"
author: "Daniel J Murphy & Paulo H Kalkhake"
date: "2 December 2016"
output: 
  ioslides_presentation:
    #css: https://maxcdn.bootstrapcdn.com/font-awesome/4.4.0/css/font-awesome.min.css
    logo: 
  beamer_presentation: default
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Set working directory
try(setwd("/Users/Paulo/GitHub/FinalPaper/presentation"),silent=TRUE)
try(setwd("/Users/djm113/Documents/GitHub/FinalPaper/presentation"),silent=TRUE)

```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- What is your research question? Why is it interesting?

- What data did you gather/methods did you use?

- What are your results?

- Engagingly present your research question and key findings to a general academic audience (fellow students).

## Estimation Results

``` {r, echo = FALSE, include=TRUE, message = FALSE, error = FALSE, results = "asis"}

analysis_simple$logABAB <- analysis_simple$AB_supply*analysis_simple$log_ABsupply

ModelIILM <- lm(log_nights ~ log_ABsupply + AB_supply + log_inc + ue_rate + marketentry + NID + factor_ym, data=analysis_simple)

ModelII <- plm(log_nights ~ log_ABsupply + AB_supply + log_inc + ue_rate + marketentry, data=analysis_simple, index=c("NID", "factor_ym"), model="within")

#Model C: introduces an interaction between log_airbnb supply and the absolute level of airbnb supply
ModelC <- plm(occup_rate ~ log_ABsupply + AB_supply + logABAB + log_inc + ue_rate, data=analysis_simple, index=c("NID", "factor_ym"), model="within")

#Model D: introduces a binary variable for Airbnb's official market entry
ModelD <- plm(occup_rate ~ log_ABsupply + AB_supply + logABAB + log_inc + ue_rate + marketentry, data=analysis_simple, index=c("NID", "factor_ym"), model="within")

stargazer(ModelIILM, ModelII, ModelC, ModelD, 
          title="Effect of Airbnb Supply on Hotels in Berlin (2010 - 2014)", 
          digits = 2, 
          covariate.labels = c("log Airbnb listings", "Airbnb listings", "Log*Level", "Average HH income (log)", "Unemployment rate"), 
          omit = c("NID", "factor_ym"),           
          omit.stat = c('f', 'ser'), # to nicely fits on the page
          omit.labels = c("Neighbourhood-specific trend", "Time trend"), 
          dep.var.labels = c("Overnight", "Occupancy Rate"),
          no.space = TRUE,
          out.header = FALSE, 
          model.names = FALSE,
          header = FALSE,
          font.size = 'small',
          add.lines = list(c('District FE?', rep('No', 4))),
          align = FALSE,
          type = "html")

```

## Slide with Plot

```{r pressure}
plot(pressure)
```

