---
title: "Airbnb vs. Hotels in Berlin: A Regression Discontinuity Approach"
author: "Daniel J Murphy & Paulo H Kalkhake"
date: "2 December 2016"
output: 
  ioslides_presentation:
    #css: https://maxcdn.bootstrapcdn.com/font-awesome/4.4.0/css/font-awesome.min.css
    logo: presentation_figures/HSG_Logo_rgb.jpg
  beamer_presentation: default
bibliography: bibliography.bib
---

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = FALSE)

# Set working directory
try(setwd("/Users/Paulo/GitHub/FinalPaper/presentation"),silent=TRUE)
try(setwd("/Users/djm113/Documents/GitHub/FinalPaper/presentation"),silent=TRUE)

# Dynamical Link to Data/Packages R script file (takes some time!)
# source("/Users/Paulo/GitHub/FinalPaper/code/finalpaper_analysis.R")

##create a theme for all graphs (courtesy of Max Wolf)
fte_theme <- function() {
      
# Generate the colors for the chart procedurally with RColorBrewer
palette <- brewer.pal("Greys", n=9)
color.background = palette[2]
color.grid.major = palette[3]
color.axis.text = palette[6]
color.axis.title = palette[7]
color.title = palette[9]
      
# Begin construction of chart
theme_bw(base_size=9) +
        
# Set the entire chart region to a light gray color
theme(panel.background=element_rect(fill=color.background, color=color.background)) +
theme(plot.background=element_rect(fill=color.background, color=color.background)) +
theme(panel.border=element_rect(color=color.background)) +
      
# Format the grid
theme(panel.grid.major=element_line(color=color.grid.major,size=.25)) +
theme(panel.grid.minor=element_blank()) +
theme(axis.ticks=element_blank()) +
      
# Format the legend, but hide by default
theme(legend.position="none") +
theme(legend.background = element_rect(fill=color.background)) +
theme(legend.text = element_text(size=7,color=color.axis.title)) +
      
# Set title and axis labels, and format these and tick marks
theme(plot.title=element_text(color=color.title, size=10, vjust=1.25)) +
theme(axis.text.x=element_text(size=7,color=color.axis.text)) +
theme(axis.text.y=element_text(size=7,color=color.axis.text)) +
theme(axis.title.x=element_text(size=8,color=color.axis.title, vjust=0)) +
theme(axis.title.y=element_text(size=8,color=color.axis.title, vjust=1.25)) +
      
# Plot margins
theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}

```

## Airbnb in Berlin

- 20,000+ Berliners have hosted travelers via Airbnb in 2015 [@Airbnb:2015aa]. 

- And more than half a million guests.

- *Rents and Zweckentfremdungsverbot:* Schafer et al found that 5,555 residential apartments in Berlin are currently being misused for Airbnb. 


Graphic?

## Previous Research

- Effect on rents: 

- *But What About The Hotel Industry?* Guttentag estimates that Airbnb sold about 15 million room nights in 2012. Similar to that of Fairfield Inn & Suites or InterContinental, two major hotel brands.

- *Airbnb Messes with Texas:* Zervas et al. found that a 10 percent size increase of the Airbnb market in Texas resulted in a .39 percent decrease in hotel revenue.

- Our paper contributes to this white spot in literature.

## Hypotheses

- *H1:* The higher the Airbnb supply in a given district in Berlin, the greater the negative effect on the hotel industry in that same district.

- *H2:* The Airbnb effect is likely the strongest once it has reached an undefined critical mass.

## Data & Variables

- Sources: Statistical Information System Berlin/Brandenburg (SBB), the Federal Statistical Office and the statistical offices of the Länder (FSO), and *InsideAirbnb.com*. 

- *Airbnb Supply:* cumulative sum of listed Airbnb Apts

- *Overnight Stays* in a given month and district

## Exploring Our Data

```{r pressure}
plot(pressure)
```

## Building the Model

In order to test our hypotheses, we will employ several models

> (2) Fixed Effects with Income and Unemployment as Economic Control Variables
${\log overnights}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

> (3) Fixed Effects with Income, Unemployment, and a Neighborhood Specific Time Trend
${\log overnights}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

> (4) Uses Hotel Occupancy Rate as the Dependent Variable & Introduces an Airbnb Interaction Term. 
${OccupRate}_{it} = \beta _i * \log Abb_{it} + \tau _i + \beta _j * X' + \varepsilon _{it}$

## Modifications

- $${Occupancy Rate}_{it} =  \frac {{Overnight Stays}_{it}}{{Hotel Beds}_{it}*{days}_{t}}$$

- Introducing dynamic Airbnb supply 

## First Results

``` {r, echo = FALSE, include=TRUE, message = FALSE, error = FALSE, results = "asis"}

```

## Prelimenary Conclusion

Policy Implications

> (1) Unlike Zervas et al., we did not find a significant negative effect of Airbnb on the hotel industry. 

> (2) 


