---
title: 'Berliner Hotels Don`t Have To Fear Airbnb - At Least For Now'
output: 
  html_document:
    toc: FALSE
    linestretch: 2
    toc_float: TRUE
    smooth_scroll: TRUE
    toc_depth: 2
    highlight: monochrome
css: "css/style.css"
bibliography: Bibliography.bib

---

```{r chunk_options, include=FALSE, results='asis'}
knitr::opts_chunk$set(cache=TRUE, fig.path = "images/")

try(setwd("/Users/Paulo/GitHub/FinalPaper/website"),silent=TRUE)
try(setwd("/Users/djm113/Documents/GitHub/FinalPaper/website"),silent=TRUE)

library(googleVis)
suppressPackageStartupMessages(library(googleVis))

```


``` {r, include=FALSE, results='asis'}
source("/Users/Paulo/GitHub/FinalPaper/code/finalpaper_analysis.R")
```
<br>

*Daniel Murphy & Paulo Kalkhake, Hertie School of Governance, Berlin*
<br>

***Airbnb claims to be part of the sharing economy and fostering sustainability in urban areas. However, data shows that the majority of Airbnb listings in most cities are entire homes, many of which are rented all year round - disrupting housing and hotels. Our study contributed to that  Browse the data for Berlin below, and see for yourself.***

<br>

#### Airbnb has changed the accomodation industry forever - and keeps growing 
<br>

Airbnb and other short-term rental platforms for apartments have become hugely popular, offering accommodation that is cheaper, better located and more 'authentic' than hotels. Thus, it is no surprise that Airbnb sold about 15 million room nights in 2012 making them as big as some of the mid-level hotel chains. With its growing success, the effects of Airbnb on the Berlin housing market has recently been the subject of a great deal of scrutiny.

In 2014, the city passed a law that forbade renting out entire apartments to tourists on Airbnb. This came in response to an unprecedented 56 percent rise in rents in Berlin between 2009 and 2014, for which Airbnb and other online hosting platforms received some of the blame. The city's law was widely welcomed by Berliners hoping to slow the forces of gentrification, but it was also welcomed by another Airbnb victim in Berlin; the hotel industry. 

Airbnb claims that their service is complementary to hotels. However, data shows that the majority of Airbnb listings in most cities are entire homes, many of which are rented all year round - disrupting housing and hotels. According to a recent study from Boston University [@zervas2016rise], a ten percent increase in Airbnb supply results in a 0.35 percent decrease in hotel room revenue. This is mainly a result of price competition, as the researchers found that hotels are reducing prices in an effort to stay in the race.

With that debate in mind we looked at what the data on 15,368 listed apartments between 2010 and 2014 might tell us on how hotels in Berlin are dealing with the upcoming star in the hospitality industry. Browse our results below, and see for yourself.

<br>

#### How big is Airbnb in Berlin?
<br>

Airbnb`s popularity in Berlin has exploded since it's official debut in the city in June 2011. To answer this questions we made a couple of simplifying assumptions. First, we said that an apartment is listed six month before its first guest review. We then calculated the cumulative sum per neighbourhood.


```{r, echo=FALSE}

analysis_simple$ymym <- as.yearmon(analysis_simple$year_month)

ggplot(analysis_simple, aes(x=ymym, y=AB_supply, group=neighbourhood, colour=neighbourhood)) +
    geom_line(alpha=0.7) + 
    labs(x="Months", y="Number of Airbnb Listings") + 
    scale_colour_discrete(name ="Neighbourhood") +
    scale_x_yearmon(as.yearmon(analysis_simple$ymym), format = "%Y-%m", n = 6)

```

#### But did the increasing number of available apartments in the city really have an effect on hotels? 

As you can see, there seems to be a positive relationship between hotel occupancy rate that is the share of beds being in use, and the change in Airbnb supply - even when we looked for the effect on neighbourhood level. 

```{r, echo=FALSE}
marketentry_dynamic <- subset(danalysis_simple, danalysis_simple$dmarketentry==1)
marketentry_static <- subset(analysis_simple, analysis_simple$dmarketentry==1)

ggplot(marketentry_dynamic, aes(x=log_ABsupply, y=occup_rate)) +
    geom_point(alpha=0.5, color="#c0392b") +
    geom_smooth(method = "lm", se = FALSE, color="black", fill="black") +
    labs(x="Log Number of Airbnb Listings", y="Hotel Occupance Rate")
```


#### But remember, Berlin's popularity increased significantly in the last years.
The rise of Berlin as the place to be, happened to coincide with the rise of Airbnb as the place to stay. Viewed this way, we find that a one percent increase in Airbnb supply relates to a xx percentage point decrease in hotel occupancy rates. 
 
#### So, is Airbnb hurting hotels in Berlin? Yes, but it is small. 
Our preliminary answer, is that our results suggests an inverse relationship between the change in Airbnb supply and hotel occupancy rates. However, you have to be careful when deriving causalities from these results. Moreover, our analysis excludes data from 2015 and 2016. Given the enormeous growth in both years, we expect the relationship to be much different. 

###Questions? 
Comments and feedback are encouraged, and can be sent to [Paulo Kalkhake](mailto: p.kalkhake@mpp.hertie-school.org(p.kalkhake@mpp.hertie-school.org)) or [Dan Murphy](mailto:danieljmurphy01@gmail.com).

### Interested? 

Take a look at these papers:
